The problem and promise of scale in multilayer animal social networks

### 2020-05-25 paper/manuscript-trad.Rmd
explore two alternatives

# Scale in multilayer networks: a case study

@@

# Scale in multilayer networks in practice: case study results and discussion
<!-- TODO: Add intro/general to results?-->
In general, individuals in multilayer networks became more connected as social
scale increased; however, this relationship varied by habitat type and season.
As social scale increased, mean graph strength for summer networks did not
change as the social threshold increased, but graph strength in winter networks
increased rapidly between 5 m and 50 m thresholds, followed by a plateau (Figure
\@ref(fig:socres)a). Layer relevance was stable around 0 during the summer but
increased for winter networks between 5 and 50 m social thresholds, then did not
change for any habitat types (Figure \@ref(fig:socres)b). This pattern suggests
the winter habitat layers captured most, if not all, of the potential
connections an individual would have across these networks. Layer similarity was
not very strong for any habitat type, indicating that summer and winter networks
are only weakly correlated. Furthermore, layer similarity in open habitats
showed more variation as social threshold increased compared to forage and
forest habitats (Figure \@ref(fig:socres)c). Connective redundancy was
relatively stable across social thresholds, especially after 50 m (Figure
\@ref(fig:socres)d). Taken together, these results suggest the optimal social
scale at which groups should be assigned is likely somewhere between ~20 - 100 m
based on the rate at which network metrics stabilized. For ungulates, groups are
often assigned based on 50 m threshold (Lingle 2003; Lesmeries et al. 2018;
Peignier et al. 2019), suggesting that given similar quantities of data, ~50 m
is likely an appropriate social scale for generating social networks in caribou
and similar species.

Increasing landcover resolution resulted in a decreased availability of forage,
a rare and patchy habitat, and corresponding low connectivity of individuals in
these habitats. Individuals in summer layers had consistent weak connections
(Figure \@ref(fig:lcres)a). Mean graph strength increased consistently in winter
forest and open habitats but varied in winter forage varied across landcover
resolutions with an increase at 250 m and decrease at 500 m (Figure
\@ref(fig:lcres)a). Layer relevance was stable for all habitat classes in summer
layers; layer relevance was significantly higher in winter and stable in open
and forest habitats but decreased for forage habitats (Figure
\@ref(fig:lcres)b). <!--This pattern during the winter indicates that
individuals had fewer neighbours in forage habitat habitats compared to their
overall neighbourhood across habitats, likely due to the patchy nature of
forage-->Layer similarity was low but stable for forest and open, except for a
sharp decrease in layer similarity in open at 100 m and 250 m (Figure
\@ref(fig:lcres)c). Layer similarity was not possible to calculate between
summer and winter forage because, at all landcover resolutions greater than 30
m, there were no individuals connected in the network layers. Individuals still
used forage habitats, but given the lower availability due to increasing
landcover resolution, at a much lower rate. Connective redundancy was moderately
high but decreased slightly as landcover resolution increased, indicating that
many neighbours were the same between layers, but the proportion of neighbours
in each layer slightly decreased at increasing landcover resolutions (Figure
\@ref(fig:lcres)d). Overall, these results indicate the importance of matching
landcover resolution to scale of selection as well as ensuring the landcover
product used has a resolution fine enough to detect landscape features and
habitats relevant to study species, such as important patchy and rare habitats.

As the number of observations used to generate multilayer networks increased,
individuals became more connected and metrics appeared to stabilize. Mean graph
strength was consistently weak in summer layers but varied in winter layers at
low number of observations, becoming stable at ~75 observations (Figure
\@ref(fig:nobs)a). Layer relevance in the summer was consistently low, but
winter layers increased between approximately 10  and 100  observations,
remaining stable with additional observations (Figure \@ref(fig:nobs)b). Like
coarse landcover resolution (Figure \@ref(fig:lcres)), low numbers of
observations were not sufficient for capturing individuals in all contexts and
therefore layer similarity could not be calculated below ~30 observations for
forest, approximately ~80 observations for open (Figure \@ref(fig:nobs)c). Layer
similarity was not possible to calculate for forage layers because there were no
connected individuals in summer forage. Once sufficient observations were
obtained, layer similarity showed some variability with increasing numbers of
observations for forest and was relatively stable in open layers. Connective
redundancy increased dramatically between ~10 and ~50 observations before
stabilizing with additional observations included (Figure \@ref(fig:nobs)d).
Given the observed trend of these multilayer metrics stabilizing after a certain
number of observations are included, the results suggest this sensitivity method
could be useful for determining the number of observations necessary for
sufficiently describing social dynamics across contexts.

<!-- TODO: I'm not sure this is relevant since this is the main way it
is already done, can we make this thought more relevant to picking scales or
multilayer networks in general? I think it's important for building robust
networks -->

Some multilayer network metrics were strongly influenced by the length of the
time window. Mean graph strength was weak but became slightly stronger with
increasing time window length during the summer, but the higher mean graph
strength during the winter showed a slight increase then decrease with
increasing winter length (Figure \@ref(fig:winlen)a). Layer relevance increased
for all summer layers as individuals became more connected with increasing
window lengths, but winter layers showed relatively stable but high layer
relevance (Figure \@ref(fig:winlen)b). This pattern indicates that the
proportion of neighbours that were the same across layers increased with a
longer time frame during the summer, but there were a high proportion of
neighbors that remained the same across layers during the winter, regardless of
time length. Furthermore, though the layers were generally not very similar
between the summer and winter, the length of the time window created slight
variability in this metric, particularly for forage and open habitat types
(Figure \@ref(fig:winlen)c). Connective redundancy increased marginally with
increasing window length, indicating that the proportions of neighbours that
were the same across layers, though relatively high in general, slightly
increased when time length was greater (Figure \@ref(fig:winlen)d). These
results suggest that an increasing window length, especially for the summer
season, may be resulting in less precisely defined seasons that are not clearly
capturing periods of aggregation and dissaggregation related to resource
availability.

Moving window positions defining seasons showed some dramatic changes in the
multilayer network metrics as the seasons were redefined. Graph strength
increased at time window position ~20 for summer and winter layers and, while
summer layers continued to increase, winter layers showed a decrease in graph
strength at time window position ~35 (Figure \@ref(fig:winpos)a). Layer
relevance increased across moving window positions for summer layers and
slightly decreased for winter layers (Figure \@ref(fig:winpos)b). These results
indicate increased connectivity in summer layers and slightly decreasing
connectivity at shifted winter time windows. Layer similarity was relatively
stable across moving window positions except for a sharp decline at window
position ~28 in forage and open layers (Figure \@ref(fig:winpos) c). Connective
redundancy showed a marginal increase with increasing window position indicating
individuals had more of the same neighbours in each habitat network (Figure
\@ref(fig:winpos) d). Similar to window length, changing the position of the
time window in this case results in a mismatch between expected periods of
resource seasonality and corresponding sociality, particularly for the summer,
highlighting the importance of carefully selecting temporal scale of analysis.

@@

# Perspectives to advance multilayer networks

@@

## Phenotypes and Individual fitness

Social network positions are frequently used to represent the social environment
that an individual experiences or their social phenotype. Multilayer networks
provide a way to bridge these different scales so that we can examine social
phenotypes in more holistic ways rather than looking at each scale separately as
has traditionally been done. From a monolayer network perspective, social
phenotypes are increasingly linked to important fitness for individuals of
various species, including primates [@Brent_2017; @Thompson_2019], dolphins
[@Stanton_2012; @Stanton_2011], and birds [@McDonald_2007; @Royle_2012]. Both
social scales, types of interactions, and temporal, seasons and stages of life,
have emerged as important for understanding fitness consequences in the context
of social network positions [@Almeling_2016; @Berger_2015; @Brent_2017;
@Holekamp_2011].

For instance, social scale is major consideration for social network analysis,
especially in the context of different types of relationships, such as grooming
and association, that are typically examined separately while answering a single
question [@Castles_2014; @Carter_2015; @Farine_2015a]. Furthermore, it is a
common trope that individuals must be associating to interact, and many studies
therefore assume that proximity is proxy for interacting [@Farine_2015a]. For
instance, baboons need to be in proximity to groom each other. However, animals
have certain interactions that do not require proximity such as long-distance
vocalization and olfactory signals [@Carter_2015]. With the advancement of
biologging technology, we have the ability to record social interactions that do
not require proximity, particularly long-distance communication networks with
devices like microphone arrays [@Snijders_2017]. Using relocation data, our case
study results indicate that there are different social patterns in different
habitat types, likely indicating different types of interactions. When the
social threshold is greater than approximately 50 m during the winter,
individuals demonstrate dramatically different graph strength (Figure 
\@ref(fig:socres)) in different habitat types, having higher graph strength in
forage and open habitats where they can likely see each other and still interact
with visual cues unlike in forest habitats. Varying social scales can have
different fitness repercussions for individuals. Rhesus macaques (*Macaca
mulatta*) had higher survivorship when they had fewer but stronger, more stable
social associates and grooming partners rather than simply more associates
[@Ellis_2019]. This finding highlights that different qualities of social scales
have lasting consequences. Multilayer networks provide the opportunity to
address these types of questions involving social scale with all the individual
layers of interaction integrated in a holistic way.

Temporal scale is also important for understanding the repercussions of social
phenotypes. Varying the time window length and position for caribou, we found
there was a lot of variation in the multilayer network metrics (Figure
\@ref(fig:winlen), Figure \@ref(fig:winpos)). These changes in the outcome
depending on how the time window is defined highlight how critical it is to
properly define the time window appropriately to address the questions and
hypotheses of interest. For instance, social phenotypes can change over ontogeny
in ways that can affect fitness [@Berger_2015; @Brent_2017]. Thus, it is
critical to carefully consider the temporal scale that is meaningful for the
proposed hypotheses. Further complicating the issue of selecting suitable time
scales across ontogeny, individuals progress through development at different
rates [@Tarka_2018]. This variation can make defining appropriate temporal
scales for testing the fitness implications of an individual’s social phenotype
over different stages of ontogeny even more difficult when they vary by
individual. However, considering fine temporal scales provide insight across
developmental stages that may otherwise be missed [@Turner_2017]. Defining
temporal scales is a challenge in multilayer networks just as in monolayer
networks; however, multilayer networks provide metrics that enable researchers
to look at similarities and differences in separate layers representing
different stages of ontogeny, or seasons as in our case study, which can be
difficult to detect.

Multilayer networks provide unique opportunities and challenges, linking these
various scales, for understanding the social phenotypes and environments of
individuals and their fitness outcomes. The multilayer network framework enables
researchers to examine these different social scales simultaneously rather than
individually, which is what has been done under most circumstances, but is not
very realistic. Metrics from multilayer networks provide an opportunity to
simplify statistical models. Interactions that happen side-by-side, proximity
associations, and long-distance interactions can be encapsulated in a single
multilayer network, and a single metric can be used to describe the various
social scales of an individual’s network position. This multilevel network
metric could be used to assess the relationship between an individual’s social
phenotype and fitness. It would also be a meaningful reduction of many
network metrics that could help simplify statistical analyses and deal with the
frequent correlations between social network metrics in different types of
networks [@Castles_2014]. Furthermore, multilayer networks provide a framework
to test hypotheses linking social phenotypes and fitness across temporal scales.
Because of the non-independent nature of social interactions and associations,
especially over time, it is hard to test hypotheses about social dynamics with a
suitable null control [@Farine_2017; @Proskurnikov_2017]. Using the framework of
multilayer networks, making layers of different temporal periods provides a way
to make appropriate null models to test questions about the dynamics of an
individual’s social network position over multiple aggregated time periods.

Conversely, picking the proper temporal scale is important but a challenge
because all layers in the multilayer network should reflect the same time
period, and the period should be long enough for each layer to have enough data
for robust networks [@Farine_2017]. These considerations may be limiting factors
when answering questions that require time frames that are relevant to different
individuals, for instance. In both mono- and multilayer networks, the effect of
scale on the relationship between an individual’s social network position and
fitness need to be considered based on the hypotheses being tested and
logistical constraints.

>>>>

# Methods

@@

# Results

In general, individuals in multilayer networks became more connected as social
scale increased; however, this relationship varied by habitat type and season.
As social scale increased, mean graph strength for summer networks did not
change as the social threshold increased, but graph strength in winter networks
increased rapidly between 5 m and 50 m thresholds, followed by a plateau (Figure
\@ref(fig:socres)a). Layer relevance was stable around 0 during the summer but
increased for winter networks between 5 and 50 m social thresholds, then did not
change for any habitat types (Figure \@ref(fig:socres)b). Layer similarity was
not very strong for any habitat type, indicating that summer and winter networks
are only weakly correlated. Furthermore, layer similarity in open habitats
showed more variation as social threshold increased compared to forage and
forest habitats (Figure \@ref(fig:socres)c). Connective redundancy was
relatively stable across social thresholds, especially after 50 m (Figure
\@ref(fig:socres)d).

Increasing landcover resolution resulted in a decreased availability of forage,
a rare and patchy habitat, and corresponding low connectivity of individuals in
these habitats. Individuals in summer layers had consistent weak connections
(Figure \@ref(fig:lcres)a). Mean graph strength increased consistently in winter
forest and open habitats but varied in winter forage varied across landcover
resolutions with an increase at 250 m and decrease at 500 m (Figure
\@ref(fig:lcres)a). Layer relevance was stable for all habitat classes in summer
layers; layer relevance was significantly higher in winter and stable in open
and forest habitats but decreased for forage habitats (Figure
\@ref(fig:lcres)b). <!--This pattern during the winter indicates that
individuals had fewer neighbours in forage habitat habitats compared to their
overall neighbourhood across habitats, likely due to the patchy nature of
forage-->Layer similarity was low but stable for forest and open, except for a
sharp decrease in layer similarity in open at 100 m and 250 m (Figure
\@ref(fig:lcres)c). Layer similarity was not possible to calculate between
summer and winter forage because, at all landcover resolutions greater than 30
m, there were no individuals connected in the network layers. Individuals still
used forage habitats, but given the lower availability due to increasing
landcover resolution, at a much lower rate. Connective redundancy was moderately
high but decreased slightly as landcover resolution increased, indicating that
many neighbours were the same between layers, but the proportion of neighbours
in each layer slightly decreased at increasing landcover resolutions (Figure
\@ref(fig:lcres)d).

As the number of observations used to generate multilayer networks increased,
individuals became more connected and metrics appeared to stabilize. Mean graph
strength was consistently weak in summer layers but varied in winter layers at
low number of observations, becoming stable at ~75 observations (Figure
\@ref(fig:nobs)a). Layer relevance in the summer was consistently low, but
winter layers increased between approximately 10  and 100  observations,
remaining stable with additional observations (Figure \@ref(fig:nobs)b). Like
coarse landcover resolution (Figure \@ref(fig:lcres)), low numbers of
observations were not sufficient for capturing individuals in all contexts and
therefore layer similarity could not be calculated below ~30 observations for
forest, approximately ~80 observations for open (Figure \@ref(fig:nobs)c). Layer
similarity was not possible to calculate for forage layers because there were no
connected individuals in summer forage. Once sufficient observations were
obtained, layer similarity showed some variability with increasing numbers of
observations for forest and was relatively stable in open layers. Connective
redundancy increased dramatically between ~10 and ~50 observations before
stabilizing with additional observations included (Figure \@ref(fig:nobs)d).

Some multilayer network metrics were strongly influenced by the length of the
time window. Mean graph strength was weak but became slightly stronger with
increasing time window length during the summer, but the higher mean graph
strength during the winter showed a slight increase then decrease with
increasing winter length (Figure \@ref(fig:winlen)a). Layer relevance increased
for all summer layers as individuals became more connected with increasing
window lengths, but winter layers showed relatively stable but high layer
relevance (Figure \@ref(fig:winlen)b).  Furthermore, though the layers were generally not very similar
between the summer and winter, the length of the time window created slight
variability in this metric, particularly for forage and open habitat types
(Figure \@ref(fig:winlen)c). Connective redundancy increased marginally with
increasing window length,

Moving window positions defining seasons showed some dramatic changes in the
multilayer network metrics as the seasons were redefined. Graph strength
increased at time window position ~20 for summer and winter layers and, while
summer layers continued to increase, winter layers showed a decrease in graph
strength at time window position ~35 (Figure \@ref(fig:winpos)a). Layer
relevance increased across moving window positions for summer layers and
slightly decreased for winter layers (Figure \@ref(fig:winpos)b). These results
indicate increased connectivity in summer layers and slightly decreasing
connectivity at shifted winter time windows. Layer similarity was relatively
stable across moving window positions except for a sharp decline at window
position ~28 in forage and open layers (Figure \@ref(fig:winpos) c). Connective
redundancy showed a marginal increase with increasing window position indicating
individuals had more of the same neighbours in each habitat network (Figure
\@ref(fig:winpos) d).

@@

# Discussion
<!-- Social scale -->
This pattern suggests
the winter habitat layers captured most, if not all, of the potential
connections an individual would have across these networks.

 Taken together, these results suggest the optimal social
scale at which groups should be assigned is likely somewhere between ~20 - 100 m
based on the rate at which network metrics stabilized. For ungulates, groups are
often assigned based on 50 m threshold (Lingle 2003; Lesmeries et al. 2018;
Peignier et al. 2019), suggesting that given similar quantities of data, ~50 m
is likely an appropriate social scale for generating social networks in caribou
and similar species.

<!-- spatial/lc -->
Overall, these results indicate the importance of matching
landcover resolution to scale of selection as well as ensuring the landcover
product used has a resolution fine enough to detect landscape features and
habitats relevant to study species, such as important patchy and rare habitats.

<!-- number of observations -->

Given the observed trend of these multilayer metrics stabilizing after a certain
number of observations are included, the results suggest this sensitivity method
could be useful for determining the number of observations necessary for
sufficiently describing social dynamics across contexts.

<!-- length time window -->
This pattern indicates that the
proportion of neighbours that were the same across layers increased with a
longer time frame during the summer, but there were a high proportion of
neighbors that remained the same across layers during the winter, regardless of
time length.

These
results suggest that an increasing window length, especially for the summer
season, may be resulting in less precisely defined seasons that are not clearly
capturing periods of aggregation and dissaggregation related to resource
availability.

Connective redundancy increased marginally with
increasing window length, indicating that the proportions of neighbours that
were the same across layers, though relatively high in general, slightly
increased when time length was greater (Figure \@ref(fig:winlen)d).

<!-- time window position -->
Similar to window length, changing the position of the
time window in this case results in a mismatch between expected periods of
resource seasonality and corresponding sociality, particularly for the summer,
highlighting the importance of carefully selecting temporal scale of analysis.

<!-- # Perspectives to advance multilayer networks  -->

@@

Social network positions are frequently used to represent the social environment
that an individual experiences or their social phenotype. Multilayer networks
provide a way to bridge these different scales to examine social
phenotypes in more holistic ways rather than looking at each scale separately as
has traditionally been done. From a monolayer network perspective, social
phenotypes are increasingly linked to important fitness for individuals of
various species, including primates [@Brent_2017; @Thompson_2019], dolphins
[@Stanton_2012; @Stanton_2011], and birds [@McDonald_2007; @Royle_2012]. Both
social scales, types of interactions, and temporal, seasons and stages of life,
have emerged as important for understanding fitness consequences in the context
of social network positions [@Almeling_2016; @Berger_2015; @Brent_2017;
@Holekamp_2011].

These considerations may be limiting factors
when answering questions that require time frames that are relevant to different
individuals, for instance. In both mono- and multilayer networks, the effect of
scale on the relationship between an individual’s social network position and
fitness need to be considered based on the hypotheses being tested and
logistical constraints.
